\chapter{The evolution of human communication}
\label{ch:communication}

\pagestyle{mystyle}

Human communication is uniquely sophisticated and remarkably pervasive.
In this chapter, we will dive head-first into studying its evolution, guided by the four methodological questions expounded in \cref{sec:evo-conclusion}.
First, we will consider different conceptions and definitions of communication in \cref{sec:comm:definition}. Then, we will consider how non-human animals and human children communicate, respectively, in \cref{sec:comm:phylogeny,sec:comm:ontogeny}. Lastly, we will discuss the utility of communication at length in \cref{sec:comm:function}. In doing so, we will also consider the evolution of human cooperation, the issue of the stability of communication, and the role deception plays in this.

\section{Conceptions of communication}
\label{sec:comm:definition}

There are many different ways in which organisms may communicate with each other, and indeed many different ways in which one may define communication. In any case, communication is a process necessarily involving a sender (communicator) and at least one receiver (addressee). In the literature, the terms 'communication' and 'signaling' are used interchangeably; therefore, I will consider them to be interchangeable as well.

% Some authors regard communication to inherently be a tool of persuasion, which then translates to their very definition of communication: for example, on the manipulative model of communication, communication can be taken to occur
An influential conception of communication, due to biologists \citet{DawkinsKrebs78}, maintains that communication occurs "when an animal, the actor, does something which appears to be the result of selection to influence the sense organs of another animal, the reactor, so that the reactor's behavior changes to the advantage of the actor" \citep[p.~283]{DawkinsKrebs78}.
In other words, communication happens when an animal does something to influence another animal's senses so as to change their behavior to the sender's benefit.
Embedded in this conception is an evolutionary component: the communicative behavior "appears to be the result of selection".

% % Here we see \citeauthor{DawkinsKrebs78} embedding a teleological explanation in their definition, when they speak of "the result of selection"; recall \cref{sec:teleology}.
% But more pressingly, they also include the aspect of benefits to the sender and thus some stance on the function of communication in it ("to the advantage of the [sender]").
% Because of this, I believe this definition to be insufficiently parsimonious in its assumptions. So let us instead turn to some conceptions of communication that are more neutral with regards to how communication may be used.

A similar, but slightly more general conception of communication is explicated by \citet{Freeberg19} in their discussion of communication as it relates to social cognition. They define communication as follows:
\begin{quoting}
    Communication involves an action or characteristic of one individual that influences the behaviour, behavioural tendency or physiology of at least one other individual in a fashion typically adaptive to both
    \hfill (p.~281)
\end{quoting}
Compared to \citeauthor{DawkinsKrebs78}'s definition, this definition widens the scope of communicators from animals to organisms in general: "All living things communicate" \citep[p.~281]{Freeberg19}.
Moreover, this definition includes the possibility of communicating to multiple receivers. The last notable difference between the two definitions is that on the former definition, communication necessarily benefits the sender, whereas on the latter, communication typically benefits both the sender and the receiver. This mutual benefit is an important aspect of communication, as we will see in \cref{sec:S-P08}.

This definition of communication due to \citet{Freeberg19} not only covers our case at hand -- humans speaking and listening to each other -- but also encompasses cases such as for example bacteria communicating population density through chemical signals \citep{Federle03}, squirrels producing alarm calls to alert their relatives of danger \citep{Sherman77}, and plants communicating distress through airborne sounds \citep{Khait23}.
A particular example of communication that we will return to in \cref{sec:S-P08} is that of the peacock's tail. Intuitively, it is perhaps not immediately obvious that this constitutes communication. However, one may think of it as the tail 'sending a message' to peahens: it signals the peacock's fitness to them.
Following the above definition, the peacock's tail constitutes communication because it is a characteristic that influences the behavior of peahens in a fashion adaptive to the peacock (because it improves his chances of reproducing) as well as the peahens (because it allows them to choose a quality male for mating).

Communication may be modeled in two fairly different ways, using the classical \emph{code model} of communication, and the \emph{ostensive-inferential} model of communication \citep{Scott-Phillips15-primate}.
In the former model, communication involves processes of coding and decoding messages. The coding, on the side of the sender, involves a mapping between the state of the world and a behavior (the behavior being the signal they send). Then decoding, on the side of the receiver, involves a mapping between two behaviors: the signal sent by the sender, and the subsequent response of the receiver. If the mappings are properly calibrated to each other, communication between sender and receiver can be said to have occurred \citep{Scott-Phillips15-primate}.

The code model is considered to be inadequate for accurately describing human communication, because it fails to account for the \emph{underdeterminacy of meaning}: merely taking the message at surface value, one cannot account for the meaning\footnote{I use the intuitive, colloquial interpretation of the term 'meaning', as I consider further explorations into the philosophy of language out of scope for this thesis.} that the message conveys to the sender (\citet{Scott-Phillips18-communication}; cf.~\citet{Quine60}'s views on radical translation).
The \emph{ostensive-inferential model} of communication, introduced by \citet{SperberWilson86} as part of their neo-Gricean framework of relevance theory, deals with this by taking into account the intentionality inherent in human communication.

\todo{Add example}
In the ostensive-inferential model, one may speak of a sender's \emph{informative intention}, which is their intending for the receiver to believe something. The sender's \emph{communicative intention} is then their intending for the receiver to believe that they have an informative intention. The sender may then express or convey this communicative intention to the receiver with an \emph{ostensive} behavior. If the receiver receives their communicative intention, then ostensive-inferential communication has occurred.

Currently, there is no evidence that any species other than humans communicate ostensively \citep{Scott-Phillips18-communication}. Consequently, one may conceptualize the code model and the ostensive-inferential model as capturing two different types of communication. The code model then captures the way that non-human animals communicate, and the ostensive-inferential model then captures the way that humans communicate with each other.

Finally, it is worth noting that Mercier nor Sperber ever defines what they take as the definition of communication. I will assume that they adhere to the ostensive-inferential conception of communication, since this theory was co-introduced by Sperber himself \citep{SperberWilson86} and Mercier and Sperber often refer to the principles of relevance theory in their writings \citep{MS09, MS11, Sperber10}.

% It is this ostensive-inferential model that I will consider to form the definition of \emph{human communication} throughout this thesis. When I speak of communication broadly construed, I will adhere to the definition of communication by \citet{Freeberg19}. This definition is compatible with the code model outlined by \citet{Scott-Phillips18-communication};
% \largeTodo{This is a weird comment, because it implicates maybe that the ostensive-inferential model is not compatible with the Freeberg definition. Make this implicature explicit, say something about how the Freeberg definition compares to the two models we have.}
% the code model, however, provides a level of detail that will not be necessary for our discussions of non-human animal communication.
% \largeTodo{Explicate the relation between the two models/definitions, because now it's very unclear why and how I'm using two different definitions at the same time}

\section{Communication in non-human animals}
\label{sec:comm:phylogeny}

Now we turn to the methodological question on evolutionary history, and have a look at communication in non-human animals.

Communication is used by non-human animals for a wide range of purposes, and it can be elicited by a number of different stimuli. Moreover, communicative behaviors can manifest themselves in different modalities: not only can animals communicate through vocalizations, they may also communicate through gestures or glances. For example, a chimpanzee can communicate to another chimpanzee that it wants to play with them by slapping the ground and directing their gaze at them \citep{Call07}.

One can broadly distinguish between communication in aggressive and cooperative interactions \citep{SeyfarthCheney03}. In aggressive interactions, primates may for example use communication in order to intimidate, by signaling their size and willingness to fight. This minimizes the chances of a physical altercation or fight actually happening, which minimizes the chance of injury for both the dominant and the subordinate animal.
In cooperative interactions on the other hand, where the interests of the signaler and the receiver overlap, communication can be used to alert others about predators, to coordinate foraging activities and to facilitate social interactions. For example, animals may communicate to each other "information about predators or the urgency of a predator’s approach, group movements, intergroup interactions, or the identities of individuals involved in social events" \citep[p.~168]{SeyfarthCheney03}.

In animals in general, vocalizations are most often elicited not by just one stimulus, but rather a complex combination of them. Moreover, the "history of interactions between the individuals involved" \citep[p.~151]{SeyfarthCheney03} can also play a role in eliciting vocalizations. As for the 'immediate' stimuli eliciting vocalizations, we may distinguish between sensory stimuli on the one hand and mental stimuli on the other. Sensory stimuli are to stimuli received through the external senses, such as visual, auditory and olfactory senses.
For example, if a male diana monkey sees a leopard (sensory stimulus), he will produce an alarm call, alerting his conspecifics of the presence of the leopard as well as signaling to the leopard that it has been detected (communicative behavior) \citep{Zuberbuhler97}.
Mental stimuli on the other hand can be viewed as the mental states one attributes to another individual.
The latter type of stimulus elicits the majority of vocalizations in human conversation, but there is no evidence that the attribution of mental states to others causes vocalizations in other animals \citep{SeyfarthCheney03}.
Moreover, there is a large debate among comparative psychologists and philosophers about whether non-human animals possess the ability to attribute mental states to others (i.e., have a theory of mind) at all; see \citet[Chapter~6]{Andrews15} for discussion.

As we will discuss deception in communication a great deal (e.g. in \cref{sec:deception}), let us briefly touch on deception in non-human animals as well.
Primates possess the ability to deceive others \todo{Elaborate};
however, compared to humans, their ability to deceive is rather primitive \citep{Dor17}.
One argument for this is that more sophisticated deception requires abilities for theory of mind -- as \citep{Lee13} puts it, "Lying in essence is ToM in action" (p.~91)" -- and primates do not possess these abilities to the extent that humans do \citep[cf.][Chapter~6]{Andrews15}.
The difference between human and primate deception also lies in the fact that language uniquely enables humans to communicate with each others' imaginations \citep{Dor17}. This opens up countless avenues for deception by the fabrication of stories. For primates on the other hand, without language, it is difficult to fabricate stories to deceive others. They can hide information: for example, they might suppress the food call they would usually expel upon finding food in order to keep this food for themselves. However, their ability to fabricate information is very limited \citep{Dor17}.

\section{Communication in children's development}
\label{sec:comm:ontogeny}

Let us now turn our attention from non-human animals to human children and consider how children communicate throughout their development.
Children first start communicating ostensively through pointing gestures at around 9--10 months of age \citep[pp.~20--21 and references therein]{Carpenter98}.
Although at first glance, pointing may seem like a simple behavior, it may be used in a number of communicative contexts to convey a wide range of messages and intentions.
For example, infants may point at a cup to indicate that they want to drink from it (i.e., pointing to request), but they may also point to a hidden object that their parent is searching for (i.e. pointing to inform).

Pointing can serve either of two communicative motives: an \emph{imperative} motive, where the pointer requests something from someone, and a \emph{declarative} motive, where the pointer shares their experiences and emotions with someone.
This basic account of pointing motives can be extended upon by distinguishing between declaratives as \emph{expressives} (sharing attitudes and emotions) and declaratives as \emph{informatives} (providing information). Furthermore, one can conceive of imperatives as a continuum, with the underlying motive ranging on a scale from individualistic -- e.g. forcing someone to do something -- to cooperative, e.g. indirectly making a request to someone by informing them of some desire \citep{Tomasello08}.
Primates only use pointing gestures imperatively, not declaratively \citep{Gomez04, Tomasello05}.
In particular, \citet{Bullinger11} found that primates only used a pointing gesture when they benefited from this act of communication, while 25-month-old infants pointed regardless of whether they themself benefited from this action. As \citet{Tomasello09} concludes, "[the infants] could not help but be informative" (p.~17).

\citet{Tomasello08} hypothesizes that in order to communicate intentionally,
like children begin doing around their first birthday, first the skills and motivations for \emph{shared intentionality} need to be present in the infant.
Shared intentionality is the "ability to participate with others in interactions involving joint goals, intentions, and attention" \citep[p.~139]{Tomasello08}. Communicative pointing behaviors in infants emerge around the same time as skills and motivations of shared intentionality do, which according to Tomasello confirms this hypothesis of dependency between them.

Tomasello further investigates what he calls \emph{pantomiming} or \emph{iconic gestures}, which are symbolic or representational gestures. \todo{Add example}
He presents empirical evidence\todo{Add evidence} that these kinds of gestures rely heavily on convention for their meaning, and that the acquisition and usage of these conventions bears a strong resemblance to the acquisition and usage of language.
Moreover, in the early stages of language production, toddlers communicate cross-modally by combining gestures with words to form propositions \citep{Iverson05}.
In this way, the development of communicative abilities in children can be considered to start in the gestural modality, with language emerging later on \citep{Tomasello08}.

In short, infants first acquire the skills and motivations needed for shared intentionality; then they acquire the skills and motivations for communicative pointing; and then they acquire the ability to use iconic gestures and language around the same time.

Let us, like in \cref{sec:comm:phylogeny}, touch briefly upon deception in ontogeny. \citet{Lee13} notes that there are two components to lying as a speech act. On the one hand, lying is governed by intentionality, and as such it requires the ability to represent mental states of oneself and others (i.e. theory of mind). On the other hand, lying is governed by conventionality, and as such it requires knowledge about social norms against lying.
Both of these components are acquired over the course of childhood, in particular in the first four years of development \citep{Lee13}; as such, lying can be considered to be a skill that has to be acquired \citep{Meibauer18}.

\input{chapters/2-4_communication-utility.tex}

\section{Conclusion}

This chapter saw us examining the evolution of human communication by studying each of the methodological questions raised in \cref{sec:evo-conclusion}. We sketched an overview of different conceptions of communication, had a look at how non-human animals and human children communicate, and we considered the utility of communication at length. In doing so, we outlined different dimensions of cooperation and discussed the evolution of human cooperation. Moreover, we introduced the problem of the stability of communication, and discussed lying and deception.

From the analysis in this chapter, the view arises that the uniquely human way of communicating evolved as a result of humans' growing dependence on cooperation for survival.
We will return to this view in \cref{ch:scrutiny}; but first, let us have a detailed look at the argumentative theory of reasoning.
