\chapter*{Acknowledgements}

First and foremost, I would like to thank Karolina for her enthusiasm and guidance throughout this project. I enjoyed our discussions immensely, and I love how their open atmosphere has made me more confident in my contributions as a philosopher.

I would like to thank the ILLC staff that have inspired me in my three years at the Master of Logic; this thesis would not exist in its current form without them.
I would also like to thank Rineke Verbrugge for introducing me to logic and inspiring me to get a master's degree in it.

I want to thank Annica, Dominique, Elias, Isabella, Jasper, Jonathan, Max, Niklas, Timo and Wouter for making my time at the MoL extraordinary. Despite me being less active because of my busy life in Utrecht, I have always felt just as much part of our group as any other member, and I think that's truly special. I have especially fond memories of our hiking trip this past summer, and I hope that our annual vacations stay a tradition.

My thanks goes out to all of my virtual coworking buddies over at Focusmate, without whom this thesis would not have been possible. You all made spending days on end at my desk fun, and especially your support in the home stretch has been very special to me.

To my parents and Dorien and Gilles: thank you for always being there for me. As the last one out of the three of us to graduate, I'm very excited to kick off this new phase of life for us as a family.

To Ronneke, thank you for your endless and unconditional love and support, not to mention your incredibly helpful research- and planning-related advice. I couldn't have done this without you.
